---
title: "p8105_hw1_ym2771"
author: "miao"
date: "9/12/2019"
output: github_document
---
# Problem 1


```{r create df1}
library(tidyverse)
df1 <- tibble(
    a = rnorm(8),
    b = a > 0,
    c = letters[1:8],
    d = factor(c("a","a","a","b","b","b","c","c"), levels = c("a","b","c"))
)
mean(df1$a)
mean(df1$b)
mean(df1$c)
mean(df1$d)
```
### summary 1
as the outcome shows, character variable c and factor variable d can not be applied to mean().

```{r convert class, eval = FALSE}
as.numeric(df1$b) 
as.numeric(df1$c)
as.numeric(df1$d)

```

```{r}
as.numeric(df1$b) * df1$a
as.factor(df1$b) * df1$a
as.numeric(as.factor(df1$b)) * df1$a
```

### summary 2
Given to the result, the logical and factor var are converted to numerical var by assigning the levels of catogries with ordering number, while the character value can not be converted. 

As the error shows, the mean() can only take **numerical and logical var** as objects to do the maths, while factor and character var can not be converted to the corresponding classes in mean function.

# Problem 2
```{r create df2}
df2 <- tibble(
    x = rnorm(500),
    y = rnorm(500),
    z = (x+y) > 1,
    num_z = as.numeric(z),
    fac_z = as.factor(z)
)
```


### Description  of df2

* The size of data frame df2 above is (`r nrow(df2)`,`r ncol(df2)`)
* The mean of x is `r mean(df2$x)`, 
* The median of x is `r median(df2$x)`
* And the standard deviation of x is `r sd(df2$x)`
* The propotion of cases which (x+y) > 1 is `r mean(df2$z)`

### plotting section

```{r scatter plot}
ggplot(df2, aes(x, y, color = z)) + geom_point()
ggsave('First plot.pdf')
ggplot(df2, aes(x, y, color = num_z)) + geom_point()
ggplot(df2, aes(x, y, color = fac_z)) + geom_point()


```

### Comment on color scale

* As shown in the plot, the color scale for logical and factor variables are assigned by the categories, which means different categories have different colors. 

* While for numerical variables, color scale is assigned by the quantile of the var, which means the numerical data was cut into 5 categories according to the quantile, and the color is changed from dark to light according to the absolute number of boundary of categories changing from minimum to maximum.


